---
title: "SEZ, BEPS Action 5, EU Code of Conduct - Methodology"
author: "Frederik Heitmüller"
date: "15 Juni 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
library(readxl)
if (!require('knitr')) install.packages('knitr'); library('knitr') 
if (!require('kableExtra')) install.packages('kableExtra'); library('kableExtra') 
if (!require('magrittr')) install.packages('magrittr'); library('magrittr') 
##for the map
if (!require('sf')) install.packages('sf'); library('sf') 
if (!require('rnaturalearth')) install.packages('rnaturalearth'); library('rnaturalearth')
if (!require('rgeos')) install.packages('rgeos'); library('rgeos') 
if (!require('rnaturalearthdata')) install.packages('rnaturalearthdata'); library('rnaturalearthdata')
if (!require('ggpubr')) install.packages('ggpubr'); library('ggpubr')
if (!require('anytime')) install.packages('anytime'); library('anytime') 
if (!require('XML')) install.packages('XML'); library('XML') 
if (!require('pdftools')) install.packages('pdftools'); library('pdftools') 
if (!require('tabulizer')) install.packages('tabulizer'); library('tabulizer') 
if (!require('rJava')) install.packages('rJava'); library('rJava')
if (!require('countrycode')) install.packages('countrycode'); library('countrycode')
if (!require('circlize')) install.packages('circlize'); library('circlize')
```

The analysis proceeds in four steps:
- determining countries in scope of analysis
- using UNCTAD data to assess prevalence of special economic zones per country
- Analysing the corporate tax benefits of SEZs in Latin America and the Caribbean
- Assessing the impact 

In a first step, the geographical scope of the analysis is determined. Since the article deals principally with the Latin American region, all countries and jurisdiction belonging to the region "Latin America & Caribbean" according to the World Bank's classification are included. Countries and their regional classification are obtained from: http://databank.worldbank.org/data/download/site-content/CLASS.xls, Accessed on 20 July 2020.The sample includes 

```{r countries included}

lac_countries <- read_excel("data_raw/CLASS.xls", skip=4)
lac_countries %<>%  rename(country = Economy, iso3c=Code, region=Region)
lac_countries %<>% filter(region=="Latin America & Caribbean") %>% select(country,iso3c)
print(nrow(lac_countries))

```
countries.
In a next step, data on the prevalence of SEZs is added from the UNCTAD World Investment report. The dataset was downloaded on June 16 2020 from [linked phrase](https://unctad.org/Sections/dite_dir/docs/WIR2019/WIR19_tab21.xlsx). 
It is unclear from the dataset whether cells containing ".." are zero values or missing values (in the sense that no information could be obtained). However, since some cells contain zero values, ".." will be treated as missing.
```{r UNCTAD SEZ data, results=FALSE}

unctad <- read_csv2("data_raw/WIR19_tab21.csv",skip=2, col_names=TRUE)
unctad  %<>%  rename(country = Economya) 
unctad %<>% drop_na(country)
unctad[178,1]<-"Jamaica"
unctad[207,1]<-"Turkmenistan"
unctad[104,1]<-"Bhutan"
unctad <- countrycustommatch::countrymatch_w_list(df=unctad, columnname="country")

lac_sez <- left_join(lac_countries, unctad[c("iso3c", "Number of SEZs (established by law)","Number of single-enterprise zones (\"free points\")","Year of the first promulgation of lawc")], by="iso3c")
lac_sez  %<>%  rename(sez_no = "Number of SEZs (established by law)", fp_no ="Number of single-enterprise zones (\"free points\")", first_y= "Year of the first promulgation of lawc") 
lac_sez %<>% mutate_at(vars(sez_no,fp_no),as.numeric)
lac_sez %<>% mutate_at(vars(first_y),anydate)
```
A number of countries (mostly Carribean island states) are not included in the UNCTAD dataset. These are
```{r print missings}
lac_sez %>% filter(is.na(sez_no)) %>% pull(country) %>% print
```
In addition, those marked as missing values in the UNCTAD dataset are
```{r other missings}
lac_sez %>% filter(sez_no=="..") %>% pull(country) %>% print
lac_sez %<>% mutate_if(is.character, list(~na_if(., "..")))
```
For the purpose of this analysis, it is assumed that these countries do not have SEZs.
The total number of SEZs in Latin America and the Caribbean is:
```{r number of SEZs in LAC}
sum(lac_sez$sez_no, na.rm=TRUE) %>% print
```
These are distributed across
```{r number of countries with SEZs}
lac_sez %>% filter(sez_no > 0) %>% nrow() %>% print
```
countries.

##Analysis of tax benefits
The next step consists in adding data that is relevant to assess the kind of tax-benefit available. Data on the "normal" tax regime in the rest of the country is downloaded from CIAT. CIAT provides time series data on "general" tax rates and and "maximum" tax rates. Since "general" excludes surcharges which are generally payable, the "maximum" rates are used. The difference of the SEZ regime from the rest of the country should therefore be regarded as the maximum benefit. Data was downloaded from <https://ciatorg.sharepoint.com/sites/cds/_layouts/15/guestaccess.aspx?docid=0b55688f22c7f4f3c9af47402923b8128&authkey=AflZ26auh2fOQoTW5EHzhfI&e=356a6a32ce38454f94eba7398c54cfca> on 16 June 2020. 
For jurisdictions where CIAT data was not available, data from KPMG was used. It was downloaded from <https://home.kpmg/xx/en/home/services/tax/tax-tools-and-resources/tax-rates-online/corporate-tax-rates-table.html> on 22 June 2020.
Finally, tax rates for jurisdictions available in neither database were searched on the IBFD Tax Summaries database.
Data was cleaned and country codes added (see separate document).

In the next step, a table was prepared where missing values for generally applicable tax rates were replaced. Rates were manually searched from IBFD Country Tax Guides and official websites of the country in question. 
Afterwards, a database of applicable tax rates in the countries' SEZ was compiled manually. The main source used were IBFD Country Tax Guides. Where these did not yield a clear result, the domestic law was directly consulted.
In order to create a map with the different type of regimes..
```{r classifying countries}

ciat_rates_clean <- readRDS("data_prepared/ciat_rates_clean.rds")
lac_sez <- left_join(lac_sez,ciat_rates_clean[c("2018", "iso3c")])
lac_sez %<>% rename(ciat_rate_2018 = "2018")
kpmg_rates_clean <- readRDS("data_prepared/kpmg_rates_clean.rds")
lac_sez <- left_join(lac_sez,kpmg_rates_clean[c("2018", "iso3c")])
lac_sez %<>% rename(kpmg_rate_2018 = "2018")

add_rates <- read_delim("data_created/additional_rates.csv", delim=";", col_names=TRUE)
lac_sez <- left_join(lac_sez,add_rates[c("additional_rate", "iso3c")])

## creating one vector with all generally applicable income tax rates. Privilege is given first to CIAT, then KPMG, then manual
lac_sez %<>% mutate(cdt_rate=coalesce(ciat_rate_2018, kpmg_rate_2018, additional_rate))

sez_rates <- read_delim("data_created/sez_rates.csv", delim=";", col_names=TRUE)
lac_sez <- left_join(lac_sez, sez_rates[c("sez_rate", "maximum_time", "iso3c")])

lac_sez$sez_no <- replace_na(lac_sez$sez_no, 0)

## calculate the benefit of the zones via-à-vis normal rate
lac_sez %<>% mutate(sez_benefit=as.double(sez_rate)/cdt_rate)

##then classification

cl <- c("Full exemption or long tax holiday", "Short tax holiday (10y or less)", "Important reduction (more than 50%)", "Moderate reduction (50% or less)", "Other tax benefit (gross)", "No benefit", "Jurisdiction without CIT", "No SEZs")

##First, low-tax jurisdiction. Then, 
lac_sez %<>% mutate(category = case_when(sez_no==0~cl[8],cdt_rate==0~cl[7], sez_rate=="Other tax incentive (gross)"~cl[5], sez_rate==0&maximum_time<11~cl[2],sez_rate==0~cl[1], sez_benefit==1~cl[6], sez_benefit<0.5~cl[3], sez_benefit>=0.5~cl[4])) 

## order the factors
lac_sez$category <- as.factor(lac_sez$category)
lac_sez$category <- ordered(lac_sez$category, levels=cl)

## create a table

lac_sez %>% select(category,country) %>% group_by(category) %>% group_nest() %>% kable() 

```



```{r analysis prevalence}

world <- ne_countries(scale = "medium", returnclass = "sf")
world %<>% rename(iso3c = iso_a3)
world <- left_join(world, lac_sez, by=c("iso3c"))
world_lac <- filter(world, region_wb == "Latin America & Caribbean")

colours <- c("Full exemption or long tax holiday"="#006d2c", 
          "Short tax holiday (10y or less)"="#31a354", 
          "Important reduction (more than 50%)"="#78c679", 
          "Moderate reduction (50% or less)"="#c2e699", 
          "Other tax benefit (gross)"="#ffffcc", 
          "No benefit"="#e41a1c", 
          "Jurisdiction without CIT"="#377eb8",
          "No SEZs"="#984ea3")


pl <- function(dat){
  ggplot(dat) +
  geom_sf(aes(fill = category))+
  scale_fill_manual(values=colours, guide=guide_legend(keywidth=0.7, keyheight=0.7))+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        axis.text = element_blank())
}

latam <- pl(dat=filter(world_lac, subregion=="South America"|subregion=="Central America"))+ theme(legend.position = "none")

car <- pl(dat=filter(world_lac, subregion=="Caribbean"))+ theme(legend.position = "none")

leg <- get_legend(pl(dat=world_lac))

ggarrange(latam, ggarrange(car, leg, ncol=1, nrow=2), ncol=2, nrow=1)


```

## Analysis of impact of BEPS Action 5 and Code of Conduct

Data regarding the work of the EU Code of Conduct Group can be obtained from the website of the European Council under <https://data.consilium.europa.eu/doc/document/ST-9639-2018-REV-4/en/pdf>.
The data is converted from pdf to csv following a prodcedure listed in the data_preparation document.

### Comparing data between both organizations
Now we can first compare whether the EU document lists all regimes that are listed in the OECD report.



```{r comparison EU and OECD data}

beps_regimes <- readRDS("data_prepared/beps_regimes_clean.rds")
cc_regimes <- readRDS("data_prepared/cc_regimes_clean.rds")

cc_regimes %>% filter(institution==c("OECD FHTP", "OECD FHTP + COCG")) %>% nrow %>% print
beps_regimes %>% nrow %>% print


```

The EU only lists 64 regimes that were reviewed by the OECD FHTP report, whereas in that report 309 regimes appear. This means that both datasets need to be taken into account.

### Selecting SEZ regimes
Neither OECD nor EU reports not classify directly which regimes are special economic zones, but the information is usually contained in the regime name. We filter the regimes by those which contain the a reference to a "zone", "centre", "city", "area" in them or which referred directly to Labuan, a Malaysian SEZ. Afterwards false hits were excluded, mainly where "centre" did not refer to a financial centre but to a regime for a "centre" function of a firm, such as treasury. In the case of the EU, the word "centre" was not used, but some further regimes that appeared to be SEZs were added.

```{r select BEPS Action 5 SEZ}

beps_regimes_sez <- filter(beps_regimes, grepl(paste(c("zone", "centre", "city", "area", "Labuan"), collapse="|"), regime_name, ignore.case=TRUE))

##manual removal of regimes falsely identified as SEZ
beps_regimes_sez <- beps_regimes_sez[-c(12:13, 16, 22:26, 28, 32, 33, 41, 44, 68, 71,72),] ## <https://thailand.ahk.de/en/members/member-news/news-details/overview-of-the-international-business-centre-regime> 
##https://lawphil.net/statutes/repacts/ra1999/ra_8756_1999.html
## https://georgiawealth.info/en/the-face-of-virtual-zone-in-georgia/
##https://juslaws.com/international-trade-centers-itc.php

nrow(beps_regimes_sez)
nrow(distinct(beps_regimes_sez, country))



```
As of July 2019 (when the latest FHTP review was published), the FHTP had reviewed 60 Special Economic Zones regimes in 32 countries worldwide (or was in the process of undertaking a review).

```{r select Code of Conduct SEZ}
cc_regimes_sez <- filter(cc_regimes, grepl(paste(c("zone", "city", "area", "Labuan", "The International Financial Services Centre (Dublin)", "Trieste Financial Services and Insurance centre", "Botswana International Financial Services Centre Companies (BITCC)"), collapse="|"), regime_name, ignore.case=TRUE)) 
# just the word 'centre' was not used as criteria since it produced too many errors. Instead, certain centres that seemed to be zones were included.
#filtering out regimes that rather seem to be disadvantaged area regimes
cc_regimes_sez %<>% filter(!identifier%in% c("BE008","DE006", "HR002", "HR003", "HR005", "IE007", "RO002", "GE004", "VN005")) 
nrow(cc_regimes_sez)
cc_regimes_sez %>% distinct(iso3c) %>% nrow


```
As of November 2019, the EU Code of Codunct Group reviewed 77 SEZ regimes in 46 countries.

### Production of a joint table.


For the remainder of the analyis, only the SEZ regimes located in Latin America are considered.

In the next step, a joint table of FHTP and Code of Conduct assessments of SEZ regimes in the LAC region will be produced. FHTP and Code of Conduct use different categories to describe status of regimes and outcomes of assessment. 
Therefore, codes in the assessment of the FHTP report will be recoded to display the intial assessment carried out by the FHTP and not the current status of the regime. The codes "Not harmful (amended)", "In the process of being eliminated/amended" and "In the process of being eliminated" imply that the regime has initially been assessed as harmful. Therefore, they will be recoded as such. "Out of scope" on the other hand implies that the regime is "Not harmful". 

If the field was left empty, this most likely implies that the regime is still "Under review".


```{r common assessment table}
beps_regimes_sez %<>% select(!c(regime_type, country))
beps_regimes_sez %<>% rename(assessment=status)
beps_regimes_sez %<>% add_column(institution="OECD FHTP")

cc_regimes_sez %<>% select(regime_name, assessment, iso3c, institution, rollback)
cc_regimes_sez %<>% rename(comments=rollback)

regimes_sez <- bind_rows(beps_regimes_sez, cc_regimes_sez, .id="source")
regimes_sez$source %<>% recode("1" = "OECD", "2" ="EU")

regimes_sez$assessment %<>% as_factor
regimes_sez$assessment %<>% fct_collapse("harmful" = 
                                           c("Not harmful (amended) except for the extension to new entrants between  July 01 and 1 October 201, which is harmful", 
                                             "Not harmful (amended)",
                                             "Harmful",
                                             "In the process of being amended",
                                             "Abolished",
                                             "In the process of being eliminated/amended",
                                             "Out of scope (amended)",
                                             "In the process of being eliminated"), 
                                         "not_harmful" =
                                           c("Not operational",
                                             "Disadvantaged areas regime",
                                             "Out of scope",
                                             "Not harmful",
                                             "Not harmful, but annual monitoring",
                                             "Not harmful17",
                                             "Not Harmful"),
                                         "under_review" = 
                                           c("Under review",
                                             "Not assessed, but annual monitoring",
                                             "Not assessed",
                                             "Not assessed14",
                                             "",
                                             "Not currently harmful (not yet operational), but annual monitoring" ))

regimes_sez$institution  %<>%    recode("OECD FHTP" = "OECD_FHTP")

write_csv2(regimes_sez, "data_created/regimes_sez_combined.csv")
saveRDS(regimes_sez, "data_created/regimes_sez_combined.rds")

regimes_sez_ctry <- regimes_sez %>% pivot_wider(id_cols=c("iso3c", "institution"), names_from=c("assessment", "institution"), values_from="regime_name", values_fn=list(assessment_n=median))


regimes_sez_ctry %<>%   add_column(harmfulregimes_found = NA)%<>% mutate(harmfulregimes_found= case_when(harmful_COCG=="NULL"&harmful_OECD_FHTP=="NULL"~"no", TRUE~"yes" ))
regimes_sez_ctry %<>% add_column(regime_assessed = "yes")



lac_sez <- left_join(lac_sez, regimes_sez_ctry[c("iso3c","harmful_COCG", "harmful_OECD_FHTP", "harmfulregimes_found","regime_assessed")])
lac_sez %<>% mutate(regime_assessed = str_replace_na(regime_assessed, "no"))
lac_sez %<>% mutate_at(.vars=c("harmful_COCG", "harmful_OECD_FHTP"), as.character)

##add inclusive framework and write to file
source("C:/Users/Frederik/surfdrive/RScripts/extracting Inclusive Framework jurisdictions.R")

complete <- left_join(complete, incframe[c("iso3c","Membership")])
complete %>% filter(is.na(Membership)) %>% nrow

write.csv2(arrange(complete[c("Country", "Harmful_COCG", "Harmful_OECD_FHTP", "Regimeassessed", "Harmfulregimesfound")], Harmfulregimesfound), "C:/Users/Frederik/surfdrive/RScripts/SEZs/complete.csv", fileEncoding="latin1")
```


```{r select Latin American}

harmfulregimes <- left_join(harmfulregimes, wb[c("iso3c", "region")], by="iso3c")

LACzones <- filter(zonesassessed, region=="Latin America & Caribbean")
kable(arrange(LACzones[1:5], Country)) %>%   kable_styling()
write.csv2(arrange(LACzones, Country), "C:/Users/Frederik/surfdrive/RScripts/SEZs/LACzones.csv", fileEncoding="latin1")

```


