---
title: "SEZ, BEPS Action 5, EU Code of Conduct - Methodology"
author: "Frederik Heitmüller"
date: "26 October 2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

if (!require('here')) install.packages('here'); library('here')
if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
library(readxl)
if (!require('knitr')) install.packages('knitr'); library('knitr') 
if (!require('kableExtra')) install.packages('kableExtra'); library('kableExtra') 
if (!require('magrittr')) install.packages('magrittr'); library('magrittr') 
##for the map
if (!require('sf')) install.packages('sf'); library('sf') 
if (!require('rnaturalearth')) install.packages('rnaturalearth'); library('rnaturalearth')
if (!require('rgeos')) install.packages('rgeos'); library('rgeos') 
if (!require('rnaturalearthdata')) install.packages('rnaturalearthdata'); library('rnaturalearthdata')
if (!require('ggpubr')) install.packages('ggpubr'); library('ggpubr')
if (!require('countrycode')) install.packages('countrycode'); library('countrycode')
if (!require('ggpattern')) remotes::install_github("coolbutuseless/ggpattern"); library('ggpattern')
library(lubridate)
if (!require('tmap')) install.packages('tmap'); library('tmap')
if (!require('patchwork')) install.packages('patchwork'); library('patchwork')
if (!require('UpSetR')) install.packages('UpSetR'); library('UpSetR')
```

The analysis proceeds in four steps:
* Determining countries in scope of analysis
* Using UNCTAD data to assess prevalence of special economic zones per country
* Analysing the corporate tax benefits of SEZs in Latin America and the Caribbean
* Assessing the impact of the FHTP's and the EU Code of Conduct Group's reviews on SEZs in Latin America

In a first step, the geographical scope of the analysis is determined. Since the article deals principally with the Latin American region, all countries and jurisdiction belonging to the region "Latin America & Caribbean" according to the World Bank's classification are included. Countries and their regional classification are obtained from: <http://databank.worldbank.org/data/download/site-content/CLASS.xls>, Accessed on 20 July 2020.The sample includes 

```{r countries included}

wb <- read_excel(here("data_raw/CLASS.xls"), skip=4)
wb %<>%  rename(country = Economy, iso3c=Code, region=Region)
lac_countries <- wb %>% filter(region=="Latin America & Caribbean") %>% select(country,iso3c)
nrow(lac_countries)

```
countries.
In a next step, data on the prevalence of SEZs is added from the UNCTAD World Investment report. The dataset was downloaded on June 16 2020 from <https://unctad.org/Sections/dite_dir/docs/WIR2019/WIR19_tab21.xlsx>. 
```{r UNCTAD SEZ data, results=FALSE}

unctad <- readRDS(here("data_prepared/unctad_sez.rds"))

lac_sez <- left_join(lac_countries, unctad[c("iso3c", "sez_no","fp_no","first_y")], by="iso3c")

```
A number of countries (mostly Carribean island states) are not included in the UNCTAD dataset or are presented as missing values. These are the following:
```{r print missings}
lac_sez %>% filter(is.na(sez_no)) %>% pull(country) %>% print
```

For the purpose of this analysis, it is assumed that these countries do not have SEZs.
The total number of SEZs in Latin America and the Caribbean is then:
```{r number of SEZs in LAC}
sum(lac_sez$sez_no, na.rm=TRUE) %>% print
```
These are distributed across 29 countries countries.
```{r number of countries with SEZs}
lac_sez %>% filter(sez_no > 0) %>% nrow() %>% print
```


## Analysis of tax benefits
The next step consists in adding data that is relevant to assess the kind of tax-benefit that the different SEZ regimes offer. For that purpose, both data on the "normal" tax regime applicable in the country to companies outside SEZs, as well as on the tax regime applicable in the SEZ is collected.

### Data on "normal" tax regime
Data on the "normal" tax regime in the rest of the country is downloaded from Centro Inter-Americano de Administración Tributaria (CIAT). CIAT provides time series data on "general" tax rates and and "maximum" tax rates. Since "general" excludes surcharges which are generally payable, the "maximum" rates are used. The difference of the SEZ regime from the rest of the country should therefore be regarded as the maximum benefit. Data was downloaded from the CIAT website: <https://ciatorg.sharepoint.com/sites/cds/_layouts/15/guestaccess.aspx?docid=0b55688f22c7f4f3c9af47402923b8128&authkey=AflZ26auh2fOQoTW5EHzhfI&e=356a6a32ce38454f94eba7398c54cfca> on 16 June 2020. 
For jurisdictions where CIAT data was not available, data from KPMG was used. It was downloaded from <https://home.kpmg/xx/en/home/services/tax/tax-tools-and-resources/tax-rates-online/corporate-tax-rates-table.html> on 22 June 2020.
Finally, tax rates for jurisdictions available in neither database were searched in the IBFD Country Tax Guides and on official websites of the country in question.
Data was cleaned and country codes added (see separate document on data preparation).
In the next step, a table was prepared where missing values for generally applicable tax rates were replaced.

```{r classifying countries}

ciat_rates_clean <- readRDS(here("data_prepared/ciat_rates_clean.rds"))
lac_sez <- left_join(lac_sez,ciat_rates_clean[c("2018", "iso3c")])
lac_sez %<>% rename(ciat_rate_2018 = "2018")
kpmg_rates_clean <- readRDS(here("data_prepared/kpmg_rates_clean.rds"))
lac_sez <- left_join(lac_sez,kpmg_rates_clean[c("2018", "iso3c")])
lac_sez %<>% rename(kpmg_rate_2018 = "2018")

add_rates <- read_delim(here("data_created/additional_rates.csv"), delim=";", col_names=TRUE)
lac_sez <- left_join(lac_sez,add_rates[c("additional_rate", "iso3c")])

## creating one vector with all generally applicable income tax rates. Privilege is given first to CIAT, then KPMG, then manual
lac_sez %<>% mutate(cdt_rate=coalesce(ciat_rate_2018, kpmg_rate_2018, additional_rate))

```

### Data on the SEZ benefit
Afterwards, a database of applicable tax benefits in the countries' SEZ was compiled manually.
In general, countries provide either for a full exemption, a fixed reduced rate (eg. 5%), or a reduction of the full tax rate (eg. 75%). There is an important difference between the latter two kinds of benefits:
Consider two countries A and B. Both have tax benefits in SEZs. In year 1, the generally applicable tax rate in both countries is 20%. The SEZ benefit in country A is a fixed rate of 5%. The SEZ benefit in country B is a 75% reduction of the full tax rate, which is equivalent to 5%. In year 2, both countries increase their general tax rates to 30%. In country A, the tax rate in SEZs will still be 5%. However, in country B, it will have increased to 7.5% (a 75% reduction from 30%).
In addition to or instead of these three permanent kinds of benefits, some SEZs provide for a tax holiday (a full tax exemption for a limited time period). When a tax holiday is available, its duration is recorded in the database. 
The main source used were IBFD Country Tax Guides. Where these did not yield a clear result, the domestic law was directly consulted. Refer to the file "data_created/sez_rates.csv" for links to the sources consulted and explanatory notes. In a few cases, the complexity of the tax regime had to be reduced for comparability reasons. In Haiti, for example, after an initial tax holiday of 15 years, the tax rate increases progressively to reach the full rate applicable in the rest of the country after six more years. Since this is difficult to capture and to compare with other countries, the data was coded as if the full rate applied directly after the expiry of the tax holiday.
When a country abolished a regime but provides for a "grandfathering period" (a period of usually a few years in which already existing companies can continue to enjoy the tax benefit), the regime is counted as if it existed until the end of the grandfathering period.
The sectoral scope of the regime was generally not recorded. The issue whether an SEZ benefit can be combined with other tax benefits was not recorded either.
In the future this dataset might be updated to include more data years.

```{r new rate dataset}
sez_rates <- read_delim(here("data_created/sez_rates.csv"), delim=";", col_names=TRUE)
lac_sez <- full_join(lac_sez, sez_rates, by="iso3c")
lac_sez %<>% filter(is.na(sez_or_disad_area)|sez_or_disad_area!="disad_area")
lac_sez <- unique(lac_sez)
# sez rate is coerced to double, remember that the row where sez_rate_cdt NA is #Paraguay with "Other incentive (gross)"
lac_sez %<>% mutate(sez_rate_cdt = coalesce(as.double(sez_rate), cdt_rate*perc_of_gen_CIT))
lac_sez %<>% mutate(sez_perc_cdt = coalesce(perc_of_gen_CIT, as.double(sez_rate)/cdt_rate))
lac_sez$sez_no <- replace_na(lac_sez$sez_no, 0)
#dummy variable for tax holiday yes or no
lac_sez %<>% mutate(holiday_dummy = case_when(holiday_time > 0~"Yes", TRUE~"No"))
lac_sez %<>% mutate(holiday_dummy_pattern = case_when(holiday_dummy== "Yes"~"stripe",TRUE~"none"))

#Categorizing the different regimes according to the type of benefit

cl <- c("Full exemption or long tax holiday", "Short tax holiday (10y or less)", "Important reduction (more than 50%)", "Moderate reduction (50% or less)", "Other tax benefit (gross)", "No benefit", "Jurisdiction without CIT", "No SEZs")


lac_sez %<>% mutate(category = case_when(sez_no==0~cl[8], cdt_rate==0~cl[7], is.na(sez_rate_cdt)~cl[5],sez_rate_cdt==0|holiday_time>10~cl[1], holiday_time>0&holiday_time<11~cl[2], sez_perc_cdt==1~cl[6], sez_perc_cdt<0.5~cl[3], sez_perc_cdt>=0.5~cl[4])) 
lac_sez$category <- ordered(as.factor(lac_sez$category), levels=cl)


# creating regime_id for non-existing regimes
lac_sez %<>% mutate_at(.vars="regime_id", coalesce, lac_sez$regime_id, lac_sez$iso3c)
# selecting the most recent year
lac_sez_rec <- lac_sez %>% group_by(regime_id) %>% slice_max(year) %>% ungroup
lac_sez_rec %<>% add_count(iso3c)
lac_sez_rec  %<>% mutate(composite = case_when(n>1~paste0(country.x, " (", regime_name,")"), TRUE~country.x))
# selecting the most favourable benefit for each jurisidiction, and the first in the table in case benefits are equal
lac_sez_rec_cds <- lac_sez_rec %>% group_by(iso3c) %>% slice_min(category) %>% ungroup
lac_sez_rec_cds %<>% group_by(iso3c) %>% slice_head %>% ungroup

lac_sez_rec %>% select(category,composite) %>% group_by(category) %>% group_nest() %>% kable() %>% kable_styling() 

```

```{r black and white map}
world <- ne_countries(scale = "medium", returnclass = "sf")
world %<>% rename(iso3c = iso_a3)
world <- left_join(world, lac_sez_rec_cds, by=c("iso3c"))
world_lac <- filter(world, region_wb == "Latin America & Caribbean")
world_lac <- filter(world, !is.na(country.y))

pl <- function(dat){
  tmap_style("bw")
tm_shape(dat, simplify=0.1, keep.units=TRUE)+
  tm_polygons("sez_perc_cdt", palette="Greys")+
  tm_symbols(col = "blue", size = "holiday_time", scale = .5)+
  tm_text("iso3c", size=0.5, remove.overlap = TRUE)
}

latam <- pl(dat=filter(world_lac, subregion=="South America"|subregion=="Central America"))
car <- pl(dat=filter(world_lac, subregion=="Caribbean"))+
  tm_layout(legend.show=FALSE)

tmap_arrange(latam, car)

```



```{r color map}

colours <- c("Full exemption or long tax holiday"="#006d2c", 
          "Short tax holiday (10y or less)"="#31a354", 
          "Important reduction (more than 50%)"="#78c679", 
          "Moderate reduction (50% or less)"="#c2e699", 
          "Other tax benefit (gross)"="#ffffcc", 
          "No benefit"="#e41a1c", 
          "Jurisdiction without CIT"="#377eb8",
          "No SEZs"="#984ea3")

pl_bl_pat <- function(dat){
ggplot(dat) +
 geom_sf_pattern(aes(fill=sez_perc_cdt, pattern_alpha=holiday_dummy), pattern="crosshatch", pattern_fill="white", pattern_spacing=0.03)+
  #scale_fill_manual(values=colours, guide=guide_legend(keywidth=0.7, keyheight=0.7))+
  scale_fill_gradient(low = "#FFFFFF", high="#000000")+
  scale_pattern_alpha_discrete(range = c(0,1))+
  theme_minimal()+
  theme(panel.grid = element_blank(),
       axis.text = element_blank())
}

pl_cl <- function(dat){ggplot(dat) +
  geom_sf(aes(fill = category))+
  scale_fill_manual(values=colours, guide=guide_legend(keywidth=0.7, keyheight=0.7))+
theme_minimal()+
  theme(panel.grid = element_blank(),
       axis.text = element_blank())
}
latam <- pl_cl(dat=filter(world_lac, subregion=="South America"|subregion=="Central America"))+ theme(legend.position = "none")
car <- pl_cl(dat=filter(world_lac, subregion=="Caribbean"))+ theme(legend.position = "none")
leg <- get_legend(pl_cl(dat=world_lac))

ggarrange(latam, ggarrange(car, leg, ncol=1, nrow=2), ncol=2, nrow=1)




```


## Analysis of impact of BEPS Action 5 and Code of Conduct

Data regarding the work of the Forum on Harmful Tax Practices (FHTP) can be obtained from <http://www.oecd.org/tax/beps/harmful-tax-practices-peer-review-results-on-preferential-regimes.pdf> and data of the EU Code of Conduct Group can be obtained from the website of the European Council under <https://data.consilium.europa.eu/doc/document/ST-9639-2018-REV-4/en/pdf>.
The data is converted from pdf to csv following a prodcedure listed in the data_preparation document.

### Comparing data between both organizations
The EU document also reports about regimes that have been reviewed by the FHTP. One can therefore compare the documents to find out whether it is sufficient to use the EU document as data source. For this purpoose we check whether whether the EU document lists all regimes that are listed in the OECD report.

```{r comparison EU and OECD data}

beps_regimes <- readRDS(here("data_prepared/beps_regimes_clean.rds"))
cc_regimes <- readRDS(here("data_prepared/cc_regimes_clean.rds"))

cc_regimes %>% filter(institution==c("OECD FHTP", "OECD FHTP + COCG")) %>% nrow %>% print
beps_regimes %>% nrow %>% print


```

The EU only lists 64 regimes that were reviewed by the OECD FHTP report, whereas in that report 309 regimes appear. This means that both datasets need to be taken into account.

### Selecting SEZ regimes
Neither OECD nor EU reports classify directly which regimes are special economic zones, but the information is usually contained in the regime name. We filter the regimes by those which contain the a reference to a "zone", "centre", "city", "area" in them or which referred directly to Labuan, a Malaysian SEZ. Afterwards false hits were excluded, mainly where "centre" did not refer to a financial centre but to a regime for a "centre" function of a firm, such as treasury. In the case of the EU, the word "centre" was not used initially (since too many errors were produced). However, some further regimes that based on their name and a quick check using google appeared to be SEZs were added.

```{r select BEPS Action 5 SEZ}

beps_regimes_sez <- filter(beps_regimes, grepl(paste(c("zone", "centre", "city", "area", "Labuan"), collapse="|"), regime_name, ignore.case=TRUE))

#manual removal of regimes falsely identified as SEZ
beps_regimes_sez <- beps_regimes_sez[-c(12:13, 16, 22:26, 28, 32, 33, 41, 44, 68, 71,72),] 
# Sources checked:
# <https://thailand.ahk.de/en/members/member-news/news-details/overview-of-the-international-business-centre-regime> 
#https://lawphil.net/statutes/repacts/ra1999/ra_8756_1999.html
# https://georgiawealth.info/en/the-face-of-virtual-zone-in-georgia/
#https://juslaws.com/international-trade-centers-itc.php

nrow(beps_regimes_sez)
nrow(distinct(beps_regimes_sez, country))

```
As of July 2019 (when the latest FHTP review was published), the FHTP had reviewed 60 Special Economic Zones regimes in 32 countries worldwide (or was in the process of undertaking a review).

```{r select Code of Conduct SEZ}
cc_regimes_sez <- filter(cc_regimes, grepl(paste(c("zone", "city", "area", "Labuan", "The International Financial Services Centre (Dublin)", "Trieste Financial Services and Insurance centre", "Botswana International Financial Services Centre Companies (BITCC)"), collapse="|"), regime_name, ignore.case=TRUE)) 
# just the word 'centre' was not used as criteria since it produced too many errors. Instead, certain centres that seemed to be zones were included.
#filtering out regimes that rather seem to be disadvantaged area regimes
cc_regimes_sez %<>% filter(!identifier%in% c("BE008","DE006", "HR002", "HR003", "HR005", "IE007", "RO002", "GE004", "VN005")) 
nrow(cc_regimes_sez)
cc_regimes_sez %>% distinct(iso3c) %>% nrow


```
As of November 2019, the EU Code of Codunct Group reviewed 77 SEZ regimes in 46 countries.

### Production of a joint table.

For the remainder of the analyis, only the SEZ regimes located in Latin America are considered.
In the next step, a joint table of FHTP and Code of Conduct assessments of SEZ regimes in the LAC region will be produced. FHTP and Code of Conduct use different categories to describe status of regimes and outcomes of assessment. 
Therefore, codes in the assessment of the FHTP report will be recoded to display the intial assessment carried out by the FHTP and not the current status of the regime. The codes "Not harmful (amended)", "In the process of being eliminated/amended" and "In the process of being eliminated" imply that the regime has initially been assessed as harmful. Therefore, they will be recoded as such. "Out of scope" on the other hand implies that the regime is "Not harmful". 
If the field was left empty, this most likely implies that the regime is still "Under review".


```{r common assessment table}
beps_regimes_sez %<>% select(!c(regime_type, country))
beps_regimes_sez %<>% rename(assessment=status)
beps_regimes_sez %<>% add_column(institution="OECD FHTP")
beps_regimes_sez$assessment %<>% as_factor

beps_regimes_sez %<>% mutate(reaction = fct_collapse(assessment, "change_introduced" = 
                                           c("Not harmful (amended) except for the extension to new entrants between  July 01 and 1 October 201, which is harmful", 
                                             "Not harmful (amended)",
                                            "Abolished",
                                             "Out of scope (amended)"),
                                           "in_process" =
                                             c("In the process of being amended",
                                              "In the process of being eliminated/amended", 
                                              "In the process of being eliminated"),
                                           "not_amended" =
                                           c("Harmful",
                                             "Not operational",
                                             "Disadvantaged areas regime",
                                             "Out of scope",
                                             "Not harmful",
                                             "Under review")))

cc_regimes_sez %<>% select(regime_name, assessment, iso3c, institution, rollback)
cc_regimes_sez %<>% rename(reaction=rollback)
cc_regimes_sez %<>% mutate(reaction = case_when(reaction==""~"not_amended", TRUE~"change_introduced"))

regimes_sez <- bind_rows(beps_regimes_sez, cc_regimes_sez, .id="source")
regimes_sez$source %<>% recode("1" = "OECD", "2" ="EU")

regimes_sez$assessment %<>% as_factor

regimes_sez$assessment %<>% fct_collapse("harmful" = 
                                           c("Not harmful (amended) except for the extension to new entrants between  July 01 and 1 October 201, which is harmful", 
                                             "Not harmful (amended)",
                                             "Harmful",
                                             "In the process of being amended",
                                             "Abolished",
                                             "In the process of being eliminated/amended",
                                             "Out of scope (amended)",
                                             "In the process of being eliminated"), 
                                         "not_harmful" =
                                           c("Not operational",
                                             "Disadvantaged areas regime",
                                             "Out of scope",
                                             "Not harmful",
                                             "Not harmful, but annual monitoring",
                                             "Not harmful17",
                                             "Not Harmful"),
                                         "under_review" = 
                                           c("Under review",
                                             "Not assessed, but annual monitoring",
                                             "Not assessed",
                                             "Not assessed14",
                                             "",
                                             "Not currently harmful (not yet operational), but annual monitoring" ))

regimes_sez$institution %<>% recode("OECD FHTP" = "OECD_FHTP")

regimes_sez <- left_join(regimes_sez, wb[c("iso3c", "region")], by="iso3c")

write_csv2(regimes_sez, here("data_created/regimes_sez_combined.csv"))
saveRDS(regimes_sez, here("data_created/regimes_sez_combined.rds"))

regimes_sez %>% filter(region=="Latin America & Caribbean") %>% arrange(iso3c) %>%  kable %>% kable_styling()

regimes_sez %>% filter(region=="Latin America & Caribbean") %>% nrow

```

# Combining SEZ regime data and data on reviews

The last part combines country data on SEZ benefits and data on reviews undertaken by the OECD and the Code of Conduct. 

```{r data combination}

regimes_sez_ctry <- regimes_sez %>% pivot_wider(id_cols=c("iso3c", "institution"), names_from=c("assessment", "institution"), values_from="regime_name")

regimes_sez_ctry %<>%   add_column(harmfulregimes_found = NA)%<>% mutate(harmfulregimes_found= case_when(harmful_COCG=="NULL"&harmful_OECD_FHTP=="NULL"~"no", TRUE~"yes" ))
regimes_sez_ctry %<>% add_column(regime_assessed = "yes")

lac_sez_rec_cds <- left_join(lac_sez_rec_cds, regimes_sez_ctry[c("iso3c","harmful_COCG", "harmful_OECD_FHTP", "harmfulregimes_found","regime_assessed")])
lac_sez_rec_cds %<>% mutate(regime_assessed = str_replace_na(regime_assessed, "no"))
lac_sez_rec_cds %<>% mutate_at(.vars=c("harmful_COCG", "harmful_OECD_FHTP"), as.character)

##add inclusive framework
inc_frame <- readRDS(here("data_prepared/inc_frame.rds"))

lac_sez_rec_cds <- left_join(lac_sez_rec_cds, inc_frame[c("iso3c","membership")])

# transform dummy variables to zeros and ones

lac_sez_rec_cds_dummy <- lac_sez_rec_cds
lac_sez_rec_cds_dummy %<>% mutate(harmfulregimes_found = case_when(harmfulregimes_found=="yes"~1, TRUE~0))
lac_sez_rec_cds_dummy %<>% mutate(regime_assessed = case_when(regime_assessed=="yes"~1, TRUE~0))
lac_sez_rec_cds_dummy %<>% mutate(sez_incentive = case_when(category=="No benefit"|category== "Jurisdiction without CIT"|category=="No SEZs" ~0, TRUE~1))
lac_sez_rec_cds_dummy %<>% add_column(all_countries = 1)
lac_sez_rec_cds_dummy %<>% mutate(has_sez = case_when(sez_no>0~1, TRUE~0))

upset(as.data.frame(lac_sez_rec_cds_dummy), sets=c("all_countries", "has_sez", "sez_incentive", "regime_assessed", "harmfulregimes_found"), keep.order=T, order.by= "degree", mainbar.y.label="No. of countries")



```

